\textit{Supplementary materials for:} The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events

### 2023-10-09 paper/supplement.tex
checking in changes from overleaf

\title{\textit{Supplementary materials for:} The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events}
\author[1]{Xinming Xu}
\author[2]{Ziyan Zhu}
\author[1, $\star$]{Jeremy R. Manning}
\affil[1]{Dartmouth College, Hanover, NH, USA}
\affil[2]{Peking University, Beijing, China}
\affil[$\star$]{Address correspondence to jeremy.r.manning@dartmouth.edu}

@@

@@

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp1.pdf}
    \caption{\textbf{Mean proportion of target events hit (A), precision (B), and convergence (C) as a function of number of segments watched in each storyline, in the main experiment participants' uncued and character-cued retrodictions and predictions.} Grey lines represent the least squares fits.}
    \label{fig:supp1}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp2.pdf}
    \caption{\textbf{Mean number of events hit for each event type, in the main experiment participants' uncued and character-cued retrodictions and predictions, averaged across just-watched segments.}}
    \label{fig:supp2}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp3.pdf}
    \caption{\textbf{(A)} References in each storyline in the main experiment stimulus. \textbf{(B)} Distribution of reference lags in the main experiment stimulus.}
    \label{fig:supp3}
\end{figure}

>>>>

\title{\textit{Supplementary materials for:} The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events}
\author[1]{Xinming Xu}
\author[2]{Ziyan Zhu}
\author[3]{Xueyao Zheng}
\author[1, $\star$]{Jeremy R. Manning}
\affil[1]{Dartmouth College, Hanover, NH, USA}
\affil[2]{Peking University, Beijing, China}
\affil[3]{Beijing Normal University, Beijing, China}
\affil[$\star$]{Address correspondence to jeremy.r.manning@dartmouth.edu}

% figures from the main text
\newcommand{\intro}{1}
\newcommand{\methods}{2}
\newcommand{\resultOne}{3}
\newcommand{\resultTwo}{4}
\newcommand{\resultThree}{5}
\newcommand{\resultFour}{6}
\newcommand{\resultFive}{7}
\newcommand{\metaAnalysis}{8}
\newcommand{\discussion}{9}

@@

\begin{table}[]
\caption{\textbf{Stimulus descriptions (replication experiment).}}
\begin{center}
\resizebox{\textwidth}{!}{
\begin{threeparttable}
\begin{tabular}{rrlll}
\toprule
\multirow{2}{*}{Segment} &
  \multirow{2}{*}{Duration (s)} &
  \multirow{2}{*}{Main characters} &
  \multicolumn{2}{l}{Number of events\tnote{a}} \\ \cmidrule(l){4-5} 
 &     &                                & Onscreen      & Offscreen\tnote{b} \\ \midrule
1 & 258 & Ji-Yoon, All faculty; Bill, Doodles\tnote{c}         & 14 (1) [1]      & 1        \\
2 & 40 & Joan, All faculty         & 3 (2)    & 0         \\
3 & 87  & Dean, Ji-Yoon      & 7 [1]          & 3         \\
4 & 78 & Elliot, Yasmine           & 5 [1]  & 1         \\
5 & 66  & Ji-Yoon, Yasmine           & 2 [1]             & 9 (1)         \\
6 & 212 & Bill; Bill, Dafna; Elliot, Yasmine        & 11 (2)    & 1         \\
7 & 134  & Ji-Yoon, Joan                & 5 (3)    & 0         \\
8 & 38 & Bill, Lila                & 2 & 0     \\
9 & 35 & Elliot, Yasmine                   & 1    & 2 (1)         \\
10 & 164  & Bill, Ji-Yoon                           & 8 [2]     & 5 \\
11 & 154 & Ji-Yoon, Habi, JuJu                     & 5 [1]             &   9 [1]        \\
12 & 216  & Joan; Ji-Yoon, Lila; Ji-Yoon, Bill         & 6 [1]     & 0         \\
13 & 56 & Ji-Yoon, Bill          & 2 [1] &           \\ \bottomrule
\end{tabular}
\begin{tablenotes}
\item[a] Number of partial events (see methods) shown in parentheses; number of summary events shown in square brackets.
\item[b] Offscreen events happened between the current segment and the next segment.
\item[c] Characters from different scenes are separated by ";".
\end{tablenotes}
\end{threeparttable}
}
\end{center}
\end{table}

@@

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{methods_rep}
      \caption{\textbf{Task overview (replication experiment).} Participants in our replication experiment watched segments from the television series \textit{The Chair}. They made free-form text responses to either retrodict what had happened in the previous segment, or predict what would happen in the next segment. We systematically varied whether participants watched the segments in forward or reverse chronological order, whether (or not) responses were cued using the main characters in the target segment, and which other segments participants had watched prior to making a response. For each segment, we collected several retrodiction or prediction across different experimental conditions. Experiment time is denoted along the vertical axis, storyline segments orders are indicated along the horizontal axis, and the colors denote experimental tasks (conditions).}
    \label{fig:methods_rep}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp1}
    \caption{\textbf{Mean proportion of target events hit (A), precision (B), and convergence (C) as a function of number of segments watched in each storyline, in participants' uncued and character-cued retrodictions and predictions (main experiment).} Grey lines represent the least squares fits.}
    \label{fig:supp1}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp1_rep}
    \caption{\textbf{Mean proportion of target events hit (A), precision (B), and convergence (C) as a function of number of segments watched, in participants' uncued and character-cued retrodictions and predictions (replication experiment).} Grey lines represent the least squares fits.}
    \label{fig:supp1_rep}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp2}
    \caption{\textbf{Mean number of events hit for each event type, in participants' uncued and character-cued retrodictions and predictions, averaged across just-watched segments (main experiment).}}
    \label{fig:supp2}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp2_rep}
    \caption{\textbf{Mean number of events hit for each event type, in participants' uncued and character-cued retrodictions and predictions, averaged across just-watched segments (replication experiment).}}
    \label{fig:supp2_rep}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp3}
    \caption{\textbf{(A)} All references in each storyline (main experiment). \textbf{(B)} Distribution of reference lags (main experiment).}
    \label{fig:supp3}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp3_rep}
    \caption{\textbf{(A)} All references (replication experiment). \textbf{(B)} Distribution of reference lags (replication experiment).}
    \label{fig:supp3_rep}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{results3_rep}

      \caption{\textbf{Characters' references drive participants' retrodiction performance (replication experiment).} \textbf{A. Reference rate as a function of lag.} Across all possible just-watched segments (lag 0), the bar heights denote the average proportions of events referenced in other past or future segments. \textbf{B. Difference in hit rates between all events and unreferenced events.} To highlight the effect of characters' references to past and future events on participants' retrodictions and predictions, here we display the difference in across-segment mean hit rates between all events and unreferenced events, as a function of temporal distance (lag) to the just-watched segment. \textbf{C. Hit rates for unreferenced events.} The average response hit rates for unreferenced events are displayed as a function of temporal distance to the just-watched segment. Error bars denote bootstrapped 95\% confidence intervals. Panels A--C: colors are described in the Figure~\resultTwo~caption. \textbf{D. Hit rates and counts of referenced and unreferenced events.} As a function of temporal distance to the just-watched segment, the sub-panels display the across-segment mean numbers ($x$-axes) and hit rates ($y$-axes) of referenced (red) and unreferenced (gray) events that participants hit (darker shading) or missed (lighter shading) in their uncued retrodictions (top sub-panel) and uncued predictions (bottom sub-panel). Intuitively, the widths of the rectangles at each lag denote the total number of events at each possible lag. The darker shading denotes the proportions of events that participants retrodicted or predicted, and the lighter shading denotes the proportions of events that participants ``missed'' in their responses.}

    \label{fig:result3_rep}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{results4_rep_ori}

        \caption{\textbf{Hit rates of reference-adjacenet events (replication experiment).} \textbf{A. Adjacent reference rate for unreferenced events as a function of lag.} Across all possible just-watched segments (lag 0), the bar heights denote the average proportion of unreferenced events in other past or future segments that were temporally adjacent to any referenced event. \textbf{B. Difference in hit rates between unreferenced events and remaining events.} To highlight the effect of reference adjacency on retrodiction and prediction of unreferenced events, here we display the difference in across-segment mean hit rates between unreferenced events and remaining events, as a function of temporal distance (lag) to the just-watched segment. \textbf{C. Hit rates for remaining events.} The across-segment mean response hit rates for unreferenced events that were \textit{not} temporally adjacent to any referenced events are displayed as a function of temporal distance to the just-watched segment. Error bars denote bootstrapped 95\% confidence intervals. Panels A--C: colors are described in the Figure~\resultTwo~caption. \textbf{D. Hit rates and counts of referenced, reference-adjacent, and remaining events.} As a function of temporal distance to the just-watched segment, the sub-panels display the numbers ($x$-axes) and proportions ($y$-axes) of referenced (red), reference-adjacent (purple), and remaining (gray) events that participants hit (darker shading) or missed (lighter shading) in their uncued retrodictions (top sub-panel) and uncued predictions (bottom sub-panel).}

    \label{fig:result4_rep_ori}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{results4_rep_sl}

      \caption{\textbf{Within-storyline reference-adjacent events are associated with higher hit rates (replication experiment).} \textbf{A. Illustration of annotation approach.} Rather than labeling reference-adjacent events as whether they were temporally adjacent to a referenced event using the original segment sequences, here we used within-storyline corrected reference-adjacency labels. We determined four storylines in the segments used in the replication experiment and assigned a storyline label for each event (shown as different background colors of each event). We then labeled reference-adjacent events as whether they were temporally adjacent to a referenced event within the same storyline (same background colors). \textbf{B. Adjacent reference rate for unreferenced events as a function of lag.} Across all possible just-watched segments (lag 0), the bar heights denote the average proportion of unreferenced events in other past or future segments that were temporally adjacent to any referenced event. \textbf{C. Difference in hit rates between unreferenced events and remaining events.} To highlight the effect of reference adjacency on retrodiction and prediction of unreferenced events, here we display the difference in across-segment mean hit rates between unreferenced events and remaining events, as a function of temporal distance (lag) to the just-watched segment. \textbf{D. Hit rates for remaining events.} The across-segment mean response hit rates for unreferenced events that were \textit{not} temporally adjacent to any referenced events are displayed as a function of temporal distance to the just-watched segment. Error bars denote bootstrapped 95\% confidence intervals. Panels A--C: colors are described in the Figure~\resultTwo~caption. \textbf{E. Hit rates and counts of referenced, reference-adjacent, and remaining events.} As a function of temporal distance to the just-watched segment, the sub-panels display the numbers ($x$-axes) and proportions ($y$-axes) of referenced (red), reference-adjacent (purple), and remaining (gray) events that participants hit (darker shading) or missed (lighter shading) in their uncued retrodictions (top sub-panel) and uncued predictions (bottom sub-panel).}

    \label{fig:results4_rep_sl}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{results5_rep}
      \caption{\textbf{Hit rates for referenced versus referring events (replication experiment).} \textbf{A. Referring rate as a function of lag.} Across all possible just-watched segments (lag 0), the bar heights denote the across-segment mean proportions of events containing references to events in other past or future segments in our replication experiment's stimuli. The bar colors are described in the Figure~\resultTwo~ caption. \textbf{B. Hit rates and counts of referenced, referring, and other events.} As a function of temporal distance to the just-watched segment, the sub-panels display the numbers ($x$-axes) and hit rates ($y$-axes) of referenced (red), referring (green), and other (gray) events that participants hit (darker shading) or missed (lighter shading) in their uncued retrodictions (top sub-panel) and uncued predictions (bottom sub-panel).}
    \label{fig:result5_rep}
\end{figure}
